### 2012-07-30 Doc/VRIPHYS2012/vriphys2012.tex
section 3
git-svn-id: svn+ssh://scm.gforge.inria.fr/svnroot/sofa-shells@438 7b2761c9-2758-49ef-8c29-9c9f1e1aad1b

For the bending formulation, the shell theory uses a measure of the out-of-plane-displacement $u^z(x,y)$. 
The strain tensor has additional contributions based on Kirchhoff-Love theory for thin plates:
\begin{equation}
\label{out-plane-strain}
\left[ \begin{array}[c]{c} \epsilon_{xx} \\ \epsilon_{yy}  \\ 2\epsilon_{yx} \end{array} \right] +=
\left[ \begin{array}[c]{c} -z \frac{\partial^2 u^z}{\partial x^2}  \vspace{1mm}\\ -z \frac{\partial^2 u^z}{\partial y^2}   \vspace{1mm}\\ - 2z \frac{\partial^2 u^z}{\partial xy}    \end{array} \right]
\end{equation}
where $z$ is the local coordinate along $z$-axis of the point on which the strain is measured.
\CD{This is something that is maybe wrong in our formulation... as the bending adds a contribution on the deformation tensor, maybe we should not separate the contribution from membrane and bending as we do now ($K_m$ and $K_b$ are computed separately) }

@@

In order to compute FEM-based stiffness for the shell element, we need to compute the jacobian matrix between strains and nodes' displacements.
For that we can reuse equation (\ref{bez-delta}) and apply the derivatives. For instance, for the in-plane displacement:
$$
\frac{\delta \epsilon_{xx}} {\delta \mathbf{x}_{2}} = \sum_{i=1}^{3} (\frac{\partial n_2}{\xi_i} + \frac{\partial n_6}{\xi_i} + \frac{\partial n_9}{\xi_i} + \frac{1}{3}\frac{\partial n_{10}}{\xi_i})\frac{\partial \xi_i}{\partial x}  
$$

Using the plate theory we now can compute the strain-displacement matrices
from the displacement field $u$. The Cauchy's strain tensor for the
elastic membrane is defined as:

\begin{equation}\label{eq-cauchy}
    J_m = \left[ \begin{matrix}
        \deriv{u_x}{x} \\
        \deriv{u_y}{y} \\
        \deriv{u_x}{y} + \deriv{u_y}{x}
    \end{matrix} \right]
\end{equation}

We apply Kirchhoff-Love theory for thin plates to obtain strain-displacement matrix for bending plate element:

\begin{equation}\label{eq-kirchhoff}
    J_b = \left[ \begin{matrix}
        - z \deriv{^2 u_z}{x^2} \\
        - z \deriv{^2 u_z}{y^2} \\
        - 2z \deriv{^2 u_z}{xy}
    \end{matrix} \right]
\end{equation}

@@

\begin{align}
    \label{eq-Km}
    K_m & = \iiint_V J_m^T M J_m \, \mathrm{d} V \\
    \label{eq-Kb}
    K_b & = \iiint_V J_b^T M J_b \, \mathrm{d} V
\end{align}

\CD{or... $K =  \iiint_V \left[\begin{matrix} J_m & J_b \end{matrix} \right]^T  M \left[\begin{matrix} J_m & J_b \end{matrix} \right]$. Could we test that option easily in the code, or is it a nightmare ??}

>>>>

For the bending formulation, the shell theory uses a measure of the out-of-plane-displacement $u^z(x,y)$. 
The strain tensor has additional contributions based on Kirchhoff-Love theory for thin plates:
\begin{equation}
\label{out-plane-strain}
\left[ \begin{array}[c]{c} \epsilon_{xx} \\ \epsilon_{yy}  \\ 2\epsilon_{yx} \end{array} \right] +=
\left[ \begin{array}[c]{c} -z \frac{\partial^2 u^z}{\partial x^2}  \vspace{1mm}\\ -z \frac{\partial^2 u^z}{\partial y^2}   \vspace{1mm}\\ - 2z \frac{\partial^2 u^z}{\partial xy}    \end{array} \right]
\end{equation}
%% comment
\CD{Here, if you want to keep a separate computation for bending and membrane (see the folllowing), use $=$ instead of $+=$ !!}
%%
where $z$ is the local coordinate along $z$-axis of the point on which the strain is measured.

@@

In order to compute FEM-based stiffness for the shell element, we need to compute the jacobian matrix between strains and nodes' displacements.
As we have separated the formulations for the in-plane and the out-plane displacements, it will result with the computation of two jacobian matrices (of size $3\times9$).
\begin{equation}
\delta \left[ \begin{array}[c]{c} \epsilon_{xx} \\ \epsilon_{yy}  \\ 2\epsilon_{yx} \end{array} \right] = 
\left[ \mathbf{J}_m \right] 
\left[ \begin{array}[c]{c} \delta \mathbf{x}_{1x} \\ \delta \mathbf{x}_{1y} \\ \delta \theta_{1z} \\ \vdots    \end{array} \right]
\end{equation}
where $\left[ \mathbf{J}_m \right]$ maps the membrane in-plane displacements and 
%
\begin{equation}
\delta \left[ \begin{array}[c]{c} \epsilon_{xx} \\ \epsilon_{yy}  \\ 2\epsilon_{yx} \end{array} \right] = 
\left[ \mathbf{J}_b \right] 
\left[ \begin{array}[c]{c} \delta \mathbf{x}_{1z} \\ \delta \theta_{1x} \\ \delta \theta_{1y} \\ \vdots    \end{array} \right]
\end{equation}
where $\left[ \mathbf{J}_b \right]$ maps the bending out-plane displacements.

To compute the values in matrices $\left[ \mathbf{J}_m \right]$ and $\left[ \mathbf{J}_b \right]$,  we can reuse equation (\ref{bez-delta}) and apply the spacial derivatives.
To illustrate, let's take one value of the matrix $\left[ \mathbf{J}_m \right]$  (first line, fourth column), we first take the kinematic between $\delta T_x$ and the concerned degree of freedom $ \delta \mathbf{x}_{2x}$:
$$
   \frac{ \delta T_x}{ \delta \mathbf{x}_{2x}} = (n_2 + n_6 + n_7 + \frac{1}{3} n_{10})
$$
Then, we apply the spacial derivatives to compute the variation of $\epsilon_{xx}$ when 
$$
\delta \epsilon_{xx} += \underbrace{\sum_{i=1}^{3} (\frac{\partial n_2}{\xi_i} + \frac{\partial n_6}{\xi_i} + \frac{\partial n_9}{\xi_i} + \frac{1}{3}\frac{\partial n_{10}}{\xi_i})\frac{\partial \xi_i}{\partial x}}_{\left[ \mathbf{J}_m \right](1,4)}  {\delta \mathbf{x}_{2x}}
$$

@@

\begin{align}
    \label{eq-Km}
    \mathbf{K}_m & = \iiint_V \mathbf{J}_m^T M \mathbf{J}_m \, \mathrm{d} V \\
    \label{eq-Kb}
    \mathbf{K}_b & = \iiint_V \mathbf{J}_b^T M \mathbf{J}_b \, \mathrm{d} V
\end{align}
%% comment
\CD{This is something that is maybe wrong in our formulation... as the bending adds a contribution on the deformation tensor, maybe we should not separate the contribution from membrane and bending as we do now ($K_m$ and $K_b$ are computed separately) }
%% comment
\CD{here, it would be... $\mathbf{K}_e =  \iiint_V \left[\begin{matrix} \mathbf{J}_m & \mathbf{J}_b \end{matrix} \right]^T  M \left[\begin{matrix} \mathbf{J}_m & \mathbf{J}_b \end{matrix} \right]$. Could we test that option easily in the code, or is it a nightmare ??}
%%
